### 2015-11-15 paper/paper.tex
add IEEE template

\documentclass[journal]{IEEEtran}

\usepackage[pdftex]{graphicx}
\DeclareGraphicsExtensions{.png}

\usepackage{cite}
\usepackage{setspace}
\usepackage{amsmath}

\begin{document}

\title{CSCI 580 : Monte Carlo Path Tracing}
\markboth{Thursday~December~3\textsuperscript{rd}~2015}
{December~3\textsuperscript{rd}~2015}
\author{Alexandre Sincennes~4955830681}
\maketitle

%===================================================

\section{Introduction}

\end{document}

### 2015-11-16 Code/Project_Tests/AssemblyInfo.cpp
add test project

@@ -1,0 +1,34 @@
+#include "stdafx.h"
+
+using namespace System::Reflection;
+using namespace System::Runtime::CompilerServices;
+using namespace System::Runtime::InteropServices;
+
+//
+// General Information about an assembly is controlled through the following
+// set of attributes. Change these attribute values to modify the information
+// associated with an assembly.
+//
+[assembly:AssemblyTitleAttribute("Project_Tests")];
+[assembly:AssemblyDescriptionAttribute("")];
+[assembly:AssemblyConfigurationAttribute("")];
+[assembly:AssemblyCompanyAttribute("")];
+[assembly:AssemblyProductAttribute("Project_Tests")];
+[assembly:AssemblyCopyrightAttribute("Copyright (c)  2015")];
+[assembly:AssemblyTrademarkAttribute("")];
+[assembly:AssemblyCultureAttribute("")];
+
+//
+// Version information for an assembly consists of the following four values:
+//
+//      Major Version
+//      Minor Version
+//      Build Number
+//      Revision
+//
+// You can specify all the value or you can default the Build and Revision Numbers
+// by using the '*' as shown below:
+
+[assembly:AssemblyVersionAttribute("1.0.*")];
+[assembly:ComVisible(false)];
+

### 2015-11-16 Code/Project_Tests/UnitTest.cpp

@@ -1,0 +1,76 @@
+#include "stdafx.h"
+#include <wchar.h>
+#include "rend.h"
+#include "gz.h"
+#include <math.h>
+#include <stdlib.h>
+
+#define EPSILON 0.001
+
+using namespace System;
+using namespace System::Text;
+using namespace System::Collections::Generic;
+using namespace Microsoft::VisualStudio::TestTools::UnitTesting;
+
+
+/*
+    CHANGE PROJECT FROM COMPILING TO EXE TO A LIB FILE FOR THIS TO WORK
+*/
+
+namespace Project_Tests
+{
+	[TestClass]
+	public ref class UnitTest
+	{
+	private:
+		TestContext^ testContextInstance;
+
+	public: 
+		/// <summary>
+		///Gets or sets the test context which provides
+		///information about and functionality for the current test run.
+		///</summary>
+		property Microsoft::VisualStudio::TestTools::UnitTesting::TestContext^ TestContext
+		{
+			Microsoft::VisualStudio::TestTools::UnitTesting::TestContext^ get()
+			{
+				return testContextInstance;
+			}
+			System::Void set(Microsoft::VisualStudio::TestTools::UnitTesting::TestContext^ value)
+			{
+				testContextInstance = value;
+			}
+		};
+
+		#pragma region Additional test attributes
+		//
+		//You can use the following additional attributes as you write your tests:
+		//
+		//Use ClassInitialize to run code before running the first test in the class
+		//[ClassInitialize()]
+		//static void MyClassInitialize(TestContext^ testContext) {};
+		//
+		//Use ClassCleanup to run code after all tests in a class have run
+		//[ClassCleanup()]
+		//static void MyClassCleanup() {};
+		//
+		//Use TestInitialize to run code before running each test
+		//[TestInitialize()]
+		//void MyTestInitialize() {};
+		//
+		//Use TestCleanup to run code after each test has run
+		//[TestCleanup()]
+		//void MyTestCleanup() {};
+		//
+		#pragma endregion 
+
+
+        /*
+        CHANGE PROJECT FROM COMPILING TO EXE TO A LIB FILE FOR THIS TO WORK
+        */
+		[TestMethod]
+		void SampleMethod()
+		{
+		};
+	};
+}

### 2015-11-16 Code/Project_Tests/stdafx.cpp

@@ -1,0 +1,5 @@
+// stdafx.cpp : source file that includes just the standard includes
+// Project_Tests.pch will be the pre-compiled header
+// stdafx.obj will contain the pre-compiled type information
+
+#include "stdafx.h"

### 2015-11-24 Code/Application5.cpp
world-space data structure re-architecting

@@ -173,18 +173,19 @@ GzMatrix	rotateY =
         specpower = 32;
         valueListShader[4] = (GzPointer)&specpower;
 
-        nameListShader[5]  = GZ_TEXTURE_MAP;
+        // get rid of textures
+        //nameListShader[5]  = GZ_TEXTURE_MAP;
 #if 0   /* set up null texture function or valid pointer */
-        valueListShader[5] = (GzPointer)0;
+        //valueListShader[5] = (GzPointer)0;
 #else
-        valueListShader[5] = (GzPointer)(tex_fun);	/* or use ptex_fun */
+        //valueListShader[5] = (GzPointer)(tex_fun);	/* or use ptex_fun */
 #endif
-        status |= GzPutAttribute(m_pRender, 6, nameListShader, valueListShader);
+        status |= GzPutAttribute(m_pRender, 5, nameListShader, valueListShader);
 
 
-	status |= GzPushMatrix(m_pRender, scale);  
-	status |= GzPushMatrix(m_pRender, rotateY); 
-	status |= GzPushMatrix(m_pRender, rotateX); 
+	//status |= GzPushMatrix(m_pRender, scale);  
+	//status |= GzPushMatrix(m_pRender, rotateY); 
+	//status |= GzPushMatrix(m_pRender, rotateX); 
 
 	if (status) exit(GZ_FAILURE); 
 
@@ -213,7 +214,7 @@ int Application5::Render()
 	*/ 
 	nameListTriangle[0] = GZ_POSITION; 
 	nameListTriangle[1] = GZ_NORMAL; 
-	nameListTriangle[2] = GZ_TEXTURE_INDEX;  
+	//nameListTriangle[2] = GZ_TEXTURE_INDEX;  
 
 	// I/O File open
 	FILE *infile;
@@ -234,6 +235,8 @@ int Application5::Render()
 	* Walk through the list of triangles, set color 
 	* and render each triangle 
 	*/ 
+
+    // set kd,ks,ka
 	while( fscanf(infile, "%s", dummy) == 1) { 	/* read in tri word */
 	    fscanf(infile, "%f %f %f %f %f %f %f %f", 
 		&(vertexList[0][0]), &(vertexList[0][1]),  
@@ -261,9 +264,13 @@ int Application5::Render()
 	     */ 
 	     valueListTriangle[0] = (GzPointer)vertexList; 
 		 valueListTriangle[1] = (GzPointer)normalList; 
[…]

### 2015-11-24 Code/rend.cpp

@@ -508,10 +508,6 @@ int GzPutTriangle(GzRender	*render, int numParts, GzToken *nameList, GzPointer	*
     GzCoord *n1 = (GzCoord *)malloc(sizeof(GzCoord));
     GzCoord *n2 = (GzCoord *)malloc(sizeof(GzCoord));
 
-    GzTextureIndex *tex0 = (GzTextureIndex *)malloc(sizeof(GzTextureIndex));
-    GzTextureIndex *tex1 = (GzTextureIndex *)malloc(sizeof(GzTextureIndex));
-    GzTextureIndex *tex2 = (GzTextureIndex *)malloc(sizeof(GzTextureIndex));
-
     for (int i = 0; i < numParts; i++) {
         GzToken token = nameList[i];
         GzPointer val = valueList[i];
@@ -544,19 +540,6 @@ int GzPutTriangle(GzRender	*render, int numParts, GzToken *nameList, GzPointer	*
 
             break;
         }
-        case GZ_TEXTURE_INDEX: {
-            const GzTextureIndex* tex[3];
-            tex[0] = &((GzTextureIndex*)val)[0];
-            tex[1] = &((GzTextureIndex*)val)[1];
-            tex[2] = &((GzTextureIndex*)val)[2];
-
-
-            // do perspective stuff
-            memcpy(tex0, tex[0], sizeof(GzTextureIndex));
-            memcpy(tex1, tex[1], sizeof(GzTextureIndex));
-            memcpy(tex2, tex[2], sizeof(GzTextureIndex));
-            break;
-        }
         default:
             return GZ_FAILURE;
         }
@@ -574,27 +557,10 @@ int GzPutTriangle(GzRender	*render, int numParts, GzToken *nameList, GzPointer	*
         n_w[1] = n1;
         n_w[2] = n2;
 
-        GzTextureIndex* t_w[3] = {};
-        t_w[0] = tex0;
-        t_w[1] = tex1;
-        t_w[2] = tex2;
+        sortVerticesCW(tri_w, n_w);
+        
 
-        sortVerticesCW(tri_w, n_w, t_w);
-        lee(render, tri_w, n_w, t_w);
     }
-
-    free(tri0);
-    free(tri1);
-    free(tri2);
-
-    free(n0);
-    free(n1);
-    free(n2);
-
-    free(tex0);
-    free(tex1);
-    free(tex2);
-
     return GZ_SUCCESS;
 }
 
[…]

### 2015-11-24 Code/rend.h

@@ -1,5 +1,7 @@
 #include "disp.h" /* include your own disp.h file (e.g. hw1)*/
 
+#include <vector>
+
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
 #define	DEFAULT_IM_Z	(-10.0)  /* world coords for image plane origin */
@@ -34,6 +36,23 @@ typedef struct {			/* define a renderer */
 }  GzRender;
 #endif
 
+#ifndef GZPROJECT
+#define GZPROJECT
+typedef struct {
+    GzCoord pos;
+    GzCoord normal;
+    GzColor Ka, Kd, Ks;
+} GzVertex;
+
+typedef struct {
+    GzVertex vertices[3];
+} GzTriangle;
+
+typedef struct {
+    std::vector<GzTriangle> tris;
+} GzWorldSpaceTriangles;
+#endif
+
 // Function declaration
 // HW2
 int GzNewRender(GzRender **render, GzDisplay *display);
@@ -64,7 +83,7 @@ int GzScaleMat(GzCoord scale, GzMatrix mat);
 int findGeneralLineEq(const GzCoord* tail, const GzCoord* head, GzCoord &out);
 
 //sort vertices CW
-int sortVerticesCW(const GzCoord* (&tri)[3], GzCoord* (&normals)[3], GzTextureIndex* (&tex)[3]);
+int sortVerticesCW(const GzCoord* (&tri)[3], GzCoord* (&normals)[3]);
 
 // linear expression evaluation algorithm
 int lee(GzRender *render, const GzCoord* (&tri)[3], GzCoord* (&normals)[3], GzTextureIndex* (&tex)[3]);

### 2015-11-24 Code/Application5.cpp
created world space tri list

@@ -197,8 +197,6 @@ GzMatrix	rotateY =
 
 int Application5::Render() 
 {
-	GzToken		nameListTriangle[3]; 	/* vertex attribute names */
-	GzPointer	valueListTriangle[3]; 	/* vertex attribute pointers */
 	GzCoord		vertexList[3];	/* vertex position coordinates */ 
 	GzCoord		normalList[3];	/* vertex normals */ 
 	GzTextureIndex  	uvList[3];		/* vertex texture map indices */ 
@@ -212,8 +210,6 @@ int Application5::Render()
 	/* 
 	* Tokens associated with triangle vertex values 
 	*/ 
-	nameListTriangle[0] = GZ_POSITION; 
-	nameListTriangle[1] = GZ_NORMAL; 
 	//nameListTriangle[2] = GZ_TEXTURE_INDEX;  
 
 	// I/O File open
@@ -236,7 +232,15 @@ int Application5::Render()
 	* and render each triangle 
 	*/ 
 
-    // set kd,ks,ka
+    GzWorldSpaceTriangles* wst = new GzWorldSpaceTriangles();
+
+    // set kd,ks,ka (copied from above for now)
+    /* Material property */
+    GzColor specularCoefficient = { 0.3, 0.3, 0.3 };
+    GzColor ambientCoefficient = { 0.1, 0.1, 0.1 };
+    GzColor diffuseCoefficient = { 0.7, 0.7, 0.7 };
+
+
 	while( fscanf(infile, "%s", dummy) == 1) { 	/* read in tri word */
 	    fscanf(infile, "%f %f %f %f %f %f %f %f", 
 		&(vertexList[0][0]), &(vertexList[0][1]),  
@@ -257,17 +261,34 @@ int Application5::Render()
 		&(normalList[2][2]), 
 		&(uvList[2][0]), &(uvList[2][1]) ); 
 
-	    /* 
-	     * Set the value pointers to the first vertex of the 	
-	     * triangle, then feed it to the renderer 
-	     * NOTE: this sequence matches the nameList token sequence
-	     */ 
-	     valueListTriangle[0] = (GzPointer)vertexList; 
-		 valueListTriangle[1] = (GzPointer)normalList; 
-		 //valueListTriangle[2] = (GzPointer)uvList; 
-		 //GzPutTriangle(m_pRender, 2, nameListTriangle, valueListTriangle); 
+        // untested
+        GzTriangle* tri = new GzTriangle();
[…]

### 2015-11-24 Code/rend.h

@@ -45,11 +45,11 @@ typedef struct {
 } GzVertex;
 
 typedef struct {
-    GzVertex vertices[3];
+    GzVertex* vertices[3];
 } GzTriangle;
 
 typedef struct {
-    std::vector<GzTriangle> tris;
+    std::vector<GzTriangle*> tris;
 } GzWorldSpaceTriangles;
 #endif
 

### 2015-11-24 Code/Application5.cpp
code hint for bernie

@@ -293,6 +293,8 @@ int Application5::Render()
 
     // DO RAYCAST HERE
 
+    // e.g. raycast_render(m_pDisplay, m_pRender, wst);
+
 	GzFlushDisplay2File(outfile, m_pDisplay); 	/* write out or update display to file*/
 	GzFlushDisplay2FrameBuffer(m_pFrameBuffer, m_pDisplay);	// write out or update display to frame buffer
 

### 2015-11-25 paper/paper.tex
first few paragraphs of paper

@@
 {December~3\textsuperscript{rd}~2015}
-\author{Alexandre Sincennes~4955830681}
+\author{
+	Sincennes, Alexandre
+	\texttt{sincenne@usc.edu}\\
+	\and
+ 	Peechankottil Manikandan, Sajini
+ 	\texttt{smanikan@usc.edu}\\
+  	\and
+	Guanzhou Liu, Vincent
+ 	\texttt{guanzhol@usc.edu}\\
+  	\and
+	Ramirez, Bernie
+ 	\texttt{berniera@usc.edu}\\
+
+}
 \maketitle
@@
 
+\section{Abstract}
+This will depend on our results.
+
+%===================================================
+
 \section{Introduction}
+Rendering via the use of rays, such as in the Monte Carlo path tracing approach, stands in stark contrast to the rasterization of triangles, which is the faster and more common approach to rendering, and the one used throughout the assignments of the CSCI 580 course.
+However, with path tracing it is possible to achieve photorealism by simulating the travel of photons, rather than the approximations used in renderers that make use of rasterization.
+INSERT STUFF RELATING TO WHAT WE ACCOMPLISHED.
+
+%===================================================
+
+\section{Prior Work}
+The origins of path tracing begin with ray tracing, first outlined in A.
+Appel's 1968 paper[1].
+In the ray tracing technique, rays are cast out from the camera, or eye, through the image plane, and into the scene.
+The first surface it collides with is the one rendered for that pixel.
+Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.\\
+
+A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.
 

### 2015-11-25 paper/paper.tex
sketch paper outline

@@
 \section{Abstract}
-This will depend on our results.
+This will depend on our results.\\
+I predict this paper to be around 4 pages since it's in a rather condensed format.
 
@@
 The first surface it collides with is the one rendered for that pixel.
-Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.\\
+Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.\\
 
-A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.
+MONTE CARLO PAPERS
+
+\section{The Monte Carlo Path Tracing Algorithm}
+explain the algorithmia
+
+\section{Implementation}
+Explain architectural changes.
+Explain what parts of Monte Carlo we did.
+
+\section{Results}
+Pictures, maybe some figures for different numbers of rays.
+
+%===================================================
+
+% SAMPLE BIBLIO
+
+SAMPLE BIBLIOGRAPHY
+
+\begin{thebibliography}{4}
+
+\bibitem{Pottinger}
+Pottinger, D.C. 2005 Terrain Analysis in Realtime Strategy Games, \emph{Ensemble Studios}.
+\\
+\bibitem{Johan}
+Hagelb\"ack, J., and Johansson, S. J. 2008. Using multiagent
+potential fields in real-time strategy games. In
+Padgham, L., and Parkes, D., eds., \emph{Proceedings of the Seventh
+International Conference on Autonomous Agents and
+Multi-agent Systems (AAMAS)}.
+\\
+\bibitem{Thurau}
+Thurau, C., and Bauckhage C. 2004. Learning human-like Movement Behavior for
+Computer Games. \emph{Gerhard Sagerer Applied Computer Science,
+Bielefeld University}.
+\\
+\bibitem{Johan_2}
+Hagelb\"ack, J., and Johansson, S.
+J.
+2008.
+The Rise of Potential Fields in Real Time Strategy Bots.
+In \emph{ Proceedings of the 4th Conference on Artificial Intelligence and Interactive Digital Entertainment (AIIDE)}.
+\end{thebibliography}
+
+%===================================================
 

### 2015-11-26 Code/Application5.cpp
Added RayCast Render.
Changed camera position.
Increased distance between camera position and image plane.

@@ -101,7 +101,7 @@ GzMatrix	rotateY =
 	0.0,	0.0,	0.0,	1.0 
 }; 
 
-#if 1 	/* set up app-defined camera if desired, else use camera defaults */
+#if 0 	/* set up app-defined camera if desired, else use camera defaults */
     camera.position[X] = -3;
     camera.position[Y] = -25;
     camera.position[Z] = -4;
@@ -267,21 +267,21 @@ int Application5::Render()
         memcpy(v0->Ka, ambientCoefficient, sizeof(GzColor));
         memcpy(v0->Kd, diffuseCoefficient, sizeof(GzColor));
         memcpy(v0->Ks, specularCoefficient, sizeof(GzColor));
-        memcpy(v0->normal, vertexList[0], sizeof(GzCoord));
+        memcpy(v0->pos, vertexList[0], sizeof(GzCoord));
         memcpy(v0->normal, normalList[0], sizeof(GzCoord));
 
         GzVertex* v1 = new GzVertex();
         memcpy(v1->Ka, ambientCoefficient, sizeof(GzColor));
         memcpy(v1->Kd, diffuseCoefficient, sizeof(GzColor));
         memcpy(v1->Ks, specularCoefficient, sizeof(GzColor));
-        memcpy(v1->normal, vertexList[1], sizeof(GzCoord));
+        memcpy(v1->pos, vertexList[1], sizeof(GzCoord));
         memcpy(v1->normal, normalList[1], sizeof(GzCoord));
 
         GzVertex* v2 = new GzVertex();
         memcpy(v2->Ka, ambientCoefficient, sizeof(GzColor));
         memcpy(v2->Kd, diffuseCoefficient, sizeof(GzColor));
         memcpy(v2->Ks, specularCoefficient, sizeof(GzColor));
-        memcpy(v2->normal, vertexList[2], sizeof(GzCoord));
+        memcpy(v2->pos, vertexList[2], sizeof(GzCoord));
         memcpy(v2->normal, normalList[2], sizeof(GzCoord));
 
         tri->vertices[0] = v0;
@@ -293,7 +293,8 @@ int Application5::Render()
 
     // DO RAYCAST HERE
 
-    // e.g. raycast_render(m_pDisplay, m_pRender, wst);
+    // e.g.
+	raycast_render(m_pRender, wst);
 
 	GzFlushDisplay2File(outfile, m_pDisplay); 	/* write out or update display to file*/
[…]

### 2015-11-26 Code/Gz.h

@@ -88,6 +88,40 @@ typedef struct  GzInput
 } GzInput;
 #endif
 
+#ifndef GZRAY
+#define GZRAY
+struct  GzRay
+{
+	GzCoord         position;			/* ray position*/
+	GzCoord			direction;			/* ray direction */
+
+	GzRay()
+		//: position(GzCoord(0.0, 0.0, 0.0)), direction(GzCoord(0.0, 0.0, 0.0)) 
+	{
+		position[0] = 0.0f;
+		position[1] = 0.0f;
+		position[2] = 0.0f;
+
+		direction[0] = 0.0f;
+		direction[1] = 0.0f;
+		direction[2] = 0.0f;
+	}
+	GzRay(const GzCoord & npos, const GzCoord & ndirection)
+		//: position(npos), direction(ndirection) 
+	{
+		position[0] = npos[0];
+		position[1] = npos[1];
+		position[2] = npos[2];
+
+		direction[0] = ndirection[0];
+		direction[1] = ndirection[1];
+		direction[2] = ndirection[2];
+	}
+
+};
+
+#endif
+
 #define RED     0         /* array indicies for color vector */
 #define GREEN   1
 #define BLUE    2

### 2015-11-26 Code/rend.cpp

@@ -913,3 +913,170 @@ int lee(GzRender *render, const GzCoord* (&tri)[3], GzCoord* (&normals)[3], GzTe
     free(tri2_color);
     return GZ_SUCCESS;
 }
+//https://en.wikipedia.org/wiki/M%C3%B6ller%E2%80%93Trumbore_intersection_algorithm
+bool rayTriangleIntersect(
+	GzCoord &orig, GzCoord &dir,
+	GzCoord &v0, const GzCoord &v1, const GzCoord &v2,
+	float &out)
+{
+	GzCoord e1, e2;  //Edge1, Edge2
+	GzCoord P, Q, T;
+	float det, inv_det, u, v;
+	float t;
+
+	//Find vectors for two edges sharing V1
+	e1[0] = v1[0] - v0[0];
+	e1[1] = v1[1] - v0[1];
+	e1[2] = v1[2] - v0[2];
+
+	e2[0] = v2[0] - v0[0];
+	e2[1] = v2[1] - v0[1];
+	e2[2] = v2[2] - v0[2];
+
+	//Begin calculating determinant - also used to calculate u parameter
+	crossProduct(dir, e2, &P);
+	//CROSS(P, D, e2);
+
+	//if determinant is near zero, ray lies in plane of triangle
+	det = dotProduct(e1, P);
+
+	//NOT CULLING
+	if (det > -0.000001 && det < 0.000001) return false;
+	inv_det = 1.f / det;
+
+	//calculate distance from V1 to ray origin
+	T[0] = orig[0] - v0[0];
+	T[1] = orig[1] - v0[1];
+	T[2] = orig[2] - v0[2];
+
+
+	//Calculate u parameter and test bound
+	u = dotProduct(T, P) * inv_det;
+	//The intersection lies outside of the triangle
+	if (u < 0.f || u > 1.f) return false;
+
+	//Prepare to test v parameter
+	//CROSS(Q, T, e1);
+	crossProduct(T, e1, &Q);
+
+	//Calculate V parameter and test bound
+	v = dotProduct(dir, Q) * inv_det;
+	//The intersection lies outside of the triangle
+	if (v < 0.f || u + v  > 1.f) return false;
+
+	t = dotProduct(e2, Q) * inv_det;
+
+	if (t > 0.000001) { //ray intersection
+		out = t;
+		return true;
+	}
+
+	// No hit, no win
+	return false;
+}
+
[…]

### 2015-11-26 Code/rend.h

@@ -4,9 +4,9 @@
 
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
-#define	DEFAULT_IM_Z	(-10.0)  /* world coords for image plane origin */
-#define	DEFAULT_IM_Y	(5.0)    /* default look-at point = 0,0,0 */
-#define	DEFAULT_IM_X	(-10.0)
+#define	DEFAULT_IM_Z	(-200)//(-10.0)  /* world coords for image plane origin */
+#define	DEFAULT_IM_Y	(0.0)//(5.0)    /* default look-at point = 0,0,0 */
+#define	DEFAULT_IM_X	(0.0)//(-10.0)
 
 #define	DEFAULT_AMBIENT	{0.1, 0.1, 0.1}
 #define	DEFAULT_DIFFUSE	{0.7, 0.6, 0.5}
@@ -82,13 +82,16 @@ int GzScaleMat(GzCoord scale, GzMatrix mat);
 // where out is {A,B,C}
 int findGeneralLineEq(const GzCoord* tail, const GzCoord* head, GzCoord &out);
 
-//sort vertices CW
+// sort vertices CW
 int sortVerticesCW(const GzCoord* (&tri)[3], GzCoord* (&normals)[3]);
 
 // linear expression evaluation algorithm
 int lee(GzRender *render, const GzCoord* (&tri)[3], GzCoord* (&normals)[3], GzTextureIndex* (&tex)[3]);
 
-//general line equation helper
+// Raycast Render:
+int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris);
+
+// general line equation helper
 // where out is {A,B,C}
 int findGeneralPlaneEq(const GzCoord* (&tri)[3], GzCoord &out, float &d);
 

### 2015-11-27 paper/paper.tex
actual bibliography: first 2 sources formatted

@@
 
-% SAMPLE BIBLIO
+\begin{thebibliography}{2}
 
-SAMPLE BIBLIOGRAPHY
+\bibitem{Appel}
+Appel, A. Some techniques for shading machine renderings of solids. \emph{AFIPS '68 (Spring) Proceedings of the April 30--May 2, 1968, spring joint computer conference}: pp. 37-45, 1968.
 
-\begin{thebibliography}{4}
+\bibitem {Whitted}
+ Whitted, T. An improved illumination model for shaded display. \emph{Comunications of the AMC}, 23(6):343–349, 1980.
 
-\bibitem{Pottinger}
-Pottinger, D.C. 2005 Terrain Analysis in Realtime Strategy Games, \emph{Ensemble Studios}.
-\\
-\bibitem{Johan}
-Hagelb\"ack, J., and Johansson, S. J. 2008. Using multiagent
-potential fields in real-time strategy games. In
-Padgham, L., and Parkes, D., eds., \emph{Proceedings of the Seventh
-International Conference on Autonomous Agents and
-Multi-agent Systems (AAMAS)}.
-\\
-\bibitem{Thurau}
-Thurau, C., and Bauckhage C. 2004. Learning human-like Movement Behavior for
-Computer Games. \emph{Gerhard Sagerer Applied Computer Science,
-Bielefeld University}.
-\\
-\bibitem{Johan_2}
-Hagelb\"ack, J., and Johansson, S.
-J.
-2008.
-The Rise of Potential Fields in Real Time Strategy Bots.
-In \emph{ Proceedings of the 4th Conference on Artificial Intelligence and Interactive Digital Entertainment (AIIDE)}.
 \end{thebibliography}

### 2015-11-27 paper/paper.tex
discuss rendering equation in paper + source

@@
 The first surface it collides with is the one rendered for that pixel.
-Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.\\
+Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.
 
-MONTE CARLO PAPERS
+\par
+
+Also instrumental for Monte Carlo path tracing was work in radiosity, which takes into account lighting emitted from diffuse surfaces rather than purely specular ones.
+J.
+Kajiya's 1986's paper[3] on the rendering equation brings a mathematical basis for these relationships, described by:
+
+\begin{center}
+$L_{\text{o}}(\mathbf x,\, \omega_{\text{o}},\, \lambda,\, t) \,=\, L_e(\mathbf x,\, \omega_{\text{o}},\, \lambda,\, t) \ +\, \int_\Omega f_r(\mathbf x,\, \omega_{\text{i}},\, \omega_{\text{o}},\, \lambda,\, t)\, L_{\text{i}}(\mathbf x,\, \omega_{\text{i}},\, \lambda,\, t)\, (\omega_{\text{i}}\,\cdot\,\mathbf n)\, \operatorname d \omega_{\text{i}}$
+\end{center}
+
+\begin{flushright}
+\par 
+(1)
+\end{flushright}
+
+\par
+SOME TEXT
 
@@
 
-\begin{thebibliography}{2}
+\begin{thebibliography}{3}
 
 \bibitem{Appel}
-Appel, A. Some techniques for shading machine renderings of solids. \emph{AFIPS '68 (Spring) Proceedings of the April 30--May 2, 1968, spring joint computer conference}: pp. 37-45, 1968.
+Appel, A. Some techniques for shading machine renderings of solids. \emph{AFIPS '68 (Spring) Proceedings of the April 30--May 2, 1968, spring joint computer conference}:37-45, 1968.
 
 \bibitem {Whitted}
- Whitted, T. An improved illumination model for shaded display. \emph{Comunications of the AMC}, 23(6):343–349, 1980.
+ Whitted, T. An improved illumination model for shaded display. \emph{Comunications of the AMC}, 23(6):343-349, 1980.
+
+\bibitem {Whitted}
+ Kajiya, J. The Rendering Equation. \emph{SIGGRAPH '86}, 20(4):143-150, 1986.
 

### 2015-11-27 paper/paper.tex
outline MCPT algorithm

@@
 \usepackage{amsmath}
+\usepackage{hyperref}
 
@@
 
-Also instrumental for Monte Carlo path tracing was work in radiosity, which takes into account lighting emitted from diffuse surfaces rather than purely specular ones.
+Also instrumental for Monte Carlo path tracing was work in radiosity, which takes into account lighting emitted from diffuse surfaces rather than specular reflection.
 J.
@@
 \par
-SOME TEXT
+Monte Carlo path tracing combines the techniques of reflected or refracted rays by recursively generating other rays, but also captures the lighting relations between diffuse surfaces.
+Using random numbers, it simulates the integration found the rendering equation by using the Russian Roulette technique for the generation of every new ray, i.e.
+a random number between zero and one is generated and compared to its colour coefficients.
 
 \section{The Monte Carlo Path Tracing Algorithm}
-explain the algorithmia
+The essentials of the Monte Carlo algorithm is surprisingly simple and can be summed as thus[4]:
+
+\begin{description}
+  \item[1)] Generate a ray from the camera (position x,y, horizontal and vertical fields of view u,v), passing through the image plane, with weight = 1
+  \item[2)] Find the point of intersection with the nearest surface
+  \item[3)] Randomly (Russian Roulette) decide whether the light is emitted or reflected (test against Kd or Ks)
+  \item[3A)] if emitted, return weight * Le
+  \item[3B)] if refleced, weight *= Ks, randomly scatter the ray according to the BRDF pdf, and go to step 2
+\end{description}
 
@@
 
-\begin{thebibliography}{3}
+\begin{thebibliography}{4}
 
@@
 
-\bibitem {Whitted}
+\bibitem {Kajiya}
 Kajiya, J. The Rendering Equation. \emph{SIGGRAPH '86}, 20(4):143-150, 1986.
+
+\bibitem {Chapter}
+\url{http://cs.brown.edu/courses/cs224/papers/mc_pathtracing.pdf} page 9
 

### 2015-11-27 paper/paper.tex
explain MCPT auxiliary algorithms

@@
   \item[3A)] if emitted, return weight * Le
-  \item[3B)] if refleced, weight *= Ks, randomly scatter the ray according to the BRDF pdf, and go to step 2
+  \item[3B)] if refleced, weight *= Ks, randomly scatter the ray according to its BRDF pdf, and go to step 2
 \end{description}
+
+In order for this algorithm to be implemented, other algorithms such as ray-triangle intersection and Phong shading (for emitted light) must also be implemented. 
+Additionally, a BRDF probability distribution function (pdf) must be created for each surface. A reasonable BRDF pdf would be a uniform distribution, wherein \emph{any} 
+vector (x,y,z) which is in the hemisphere of the surface is randomly selected.
 

### 2015-11-27 paper/paper.tex
fix typo
another typo fix

@@
 The first surface it collides with is the one rendered for that pixel.
-Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.
+Additionally, rays called shadow rays are cast from the light sources and determine whether or not the surface is being occluded by another, which would be collided with first by the shadow ray.
+A further development was Whitted ray tracing[2], wherein rays generate new rays upon collision with a surface, for direct illumination, perfect refraction and perfect reflection.
 
@@
 \section{The Monte Carlo Path Tracing Algorithm}
-The essentials of the Monte Carlo algorithm is surprisingly simple and can be summed as thus[4]:
+The essentials of the Monte Carlo algorithm are surprisingly simple and can be summed as thus[4]:
 

### 2015-11-27 Code/rend.cpp
RayCasting code
Not complete

@@ -9,6 +9,7 @@
 #include	<limits>
 
 #define PI 3.141592653589793f
+#define MAXREFLECTANCE 10
 
 // HW3 helpers
 float degToRad(float deg) {
@@ -1068,6 +1069,13 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 				//if (t != 0)
 				if (test)
 				{
+					//This is where I should check for ray intersection
+					//Check if the ray intersects a surface,
+					//if so fetch the colour of the surface and assign a new direction to the ray from the point of intersection
+					
+					//Create a recursive method to reflect ray and fetch colour
+					float* newColor =  TracePath(ray, 0);		
+					
 					GzPutDisplay(render->display, i, j, 4095.0f, 0.0f, 0.0f, 1.0f, 1.0f);
 				}
 			}
@@ -1080,3 +1088,36 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 }
 
 
+float* TracePath(GzRay ray, int depth)
+{
+	if (depth == MAXREFLECTANCE)
+	{
+		GzCoord black; 
+		black[0] = 0.0f; //r g b
+		black[1] = 0.0f;
+		black[2] = 0.0f;
+
+		return black; 
+	}
+
+	//Fetch the colour of the ray position 
+	GzColor c; // ray.colour(some way to assign the colour of the object)
+
+	GzRay newray;
+	newray.position[0] = ray.position[0]; // newray origin
+	newray.position[1] = ray.position[1];
+	newray.position[2] = ray.position[2];
+
+	//newray.direction = pick a random direction 
+
+
+	// Compute the BRDF for this ray (assuming Lambertian reflection)
+	
+	GzCoord normalWhereObjectWasHit; //create a normal and assign
+	float cos_theta = dotProduct(newray.direction, normalWhereObjectWasHit);
+//	float* BRDF = 2 * m.reflectance * cos_theta;*/
+	float* reflected = TracePath(newray, depth + 1);
+
+	// Apply the Rendering Equation here.
+	//return emittance + (BRDF * reflected);*/
+}

### 2015-11-29 Code/Gz.h
Added Emitted Portion of Alg.

@@ -94,6 +94,7 @@ struct  GzRay
 {
 	GzCoord         position;			/* ray position*/
 	GzCoord			direction;			/* ray direction */
+	GzColor			color;
 
 	GzRay()
 		//: position(GzCoord(0.0, 0.0, 0.0)), direction(GzCoord(0.0, 0.0, 0.0)) 

### 2015-11-29 Code/rend.cpp

@@ -1044,6 +1044,9 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			ray.direction[2] = pixelPos[2];
 			normalizeGzCoord(ray.direction);
 
+			//float* newColor = TracePath(ray, 0, tris, render, i, j);
+
+
 			//for each triangle
 			for (int k = 0; k < tris->tris.size() - 1; k++)
 			{
@@ -1062,11 +1065,42 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 				triC[1] = tris->tris[k]->vertices[2]->pos[1];
 				triC[2] = tris->tris[k]->vertices[2]->pos[2];
 
-				//double t = ray_intersect(ray, triA, triB, triC);
+				GzCoord triANorm;
+				triANorm[0] = tris->tris[k]->vertices[0]->normal[0];//(tri0[0])[0];
+				triANorm[1] = tris->tris[k]->vertices[0]->normal[1];
+				triANorm[2] = tris->tris[k]->vertices[0]->normal[2];
+
+				GzCoord triBNorm;
+				triBNorm[0] = tris->tris[k]->vertices[1]->normal[0];
+				triBNorm[1] = tris->tris[k]->vertices[1]->normal[1];
+				triBNorm[2] = tris->tris[k]->vertices[1]->normal[2];
+
+				GzCoord triCNorm;
+				triCNorm[0] = tris->tris[k]->vertices[2]->normal[0];
+				triCNorm[1] = tris->tris[k]->vertices[2]->normal[1];
+				triCNorm[2] = tris->tris[k]->vertices[2]->normal[2];
+
+				const GzCoord* tri_w[3] = {};
+				tri_w[0] = &triA;
+				tri_w[1] = &triB;
+				tri_w[2] = &triC;
+
+				GzCoord* tri_n[3] = {};
+				tri_n[0] = &triANorm;
+				tri_n[1] = &triBNorm;
+				tri_n[2] = &triCNorm;
+
+				sortVerticesCW(tri_w, tri_n);
+
+				//float* newColor = TracePath(ray, 0,tris,render,i,j);
+
+				//GzPutDisplay(render->display, i, j, 4095.0f, 0.0f, 0.0f, 1.0f, 1.0f);
+				//GzPutDisplay(render->display, i, j, ctoi(ray.color[0]), ctoi(ray.color[1]), ctoi(ray.color[2]), 1.0f, 1.0f);
+
+
[…]

### 2015-11-29 Code/rend.h

@@ -100,3 +100,5 @@ int findGeneralPlaneEq(const GzCoord* (&tri)[3], GzCoord &out, float &d);
 #define C	2
 
 void normalizeGzCoord(GzCoord &v);
+
+float* TracePath(GzRay ray, int depth, GzWorldSpaceTriangles *tris, GzRender *render,int i,int j); //Remove i and j

### 2015-12-01 Code/Application5.cpp
Fix Normals!

@@ -236,11 +236,17 @@ int Application5::Render()
 
     // set kd,ks,ka (copied from above for now)
     /* Material property */
-    GzColor specularCoefficient = { 0.3, 0.3, 0.3 };
+	GzColor specularCoefficient = { 0.3, 0.3, 0.3 }; //Using for color
     GzColor ambientCoefficient = { 0.1, 0.1, 0.1 };
-    GzColor diffuseCoefficient = { 0.7, 0.7, 0.7 };
+    GzColor diffuseCoefficient = { 0.7, 0.7, 0.7 }; //Using for emission
 
 
+
+	int left_wall = 1;
+	int right_wall = 3;
+	int back_wall = 2;
+
+	int counter = 0;
 	while( fscanf(infile, "%s", dummy) == 1) { 	/* read in tri word */
 	    fscanf(infile, "%f %f %f %f %f %f %f %f", 
 		&(vertexList[0][0]), &(vertexList[0][1]),  
@@ -260,6 +266,38 @@ int Application5::Render()
 		&(normalList[2][0]), &(normalList[2][1]), 	
 		&(normalList[2][2]), 
 		&(uvList[2][0]), &(uvList[2][1]) ); 
+
+		/*if (counter <= left_wall)
+		{
+			specularCoefficient[0] = 1;
+			specularCoefficient[1] = 0;
+			specularCoefficient[2] = 0;
+
+			diffuseCoefficient[0] = 1;
+			diffuseCoefficient[1] = 0;
+			diffuseCoefficient[2] = 0;
+
+			ambientCoefficient[0] = 1;
+			ambientCoefficient[1] = 0;
+			ambientCoefficient[2] = 0;
+			
+			
+			
+		}
+		else if (counter <= right_wall)
+		{
+			specularCoefficient[0] = 0;
+			specularCoefficient[1] = 0;
+			specularCoefficient[2] = 1;
+
+			diffuseCoefficient[0] = 0;
+			diffuseCoefficient[1] = 0;
+			diffuseCoefficient[2] = 1;
+
+			ambientCoefficient[0] = 0;
+			ambientCoefficient[1] = 0;
+			ambientCoefficient[2] = 1;
+		}*/
 
         // untested
         GzTriangle* tri = new GzTriangle();
@@ -289,6 +327,7 @@ int Application5::Render()
         tri->vertices[2] = v2;
 
         wst->tris.push_back(tri);
+		counter++;
 	} 
 
     // DO RAYCAST HERE

### 2015-12-01 Code/disp.cpp

@@ -54,7 +54,7 @@ int GzInitDisplay(GzDisplay	*display)
 /* HW1.5 set everything to some default values - start a new frame */
 	for (int i = 0; i < display->xres; i++) {
 		for (int j = 0; j < display->yres; j++) {
-			display->fbuf[i + display->xres*j] = GzPixel{ GREY, GREY, GREY, 1, INT_MAX };
+			display->fbuf[i + display->xres*j] = GzPixel{ GREY, 0, GREY, 1, INT_MAX };
 		}
 	}
 	return GZ_SUCCESS;

### 2015-12-01 Code/rend.cpp

@@ -86,10 +86,10 @@ void barycentricCoords(const GzCoord &vert0, const GzCoord &vert1, const GzCoord
     memcpy(pointz, point, sizeof(GzCoord));
 
     //reduce z range to prevent overflow
-    vert0z[Z] /= 100000;
-    vert1z[Z] /= 100000;
-    vert2z[Z] /= 100000;
-    pointz[Z] /= 100000;
+    //vert0z[Z] /= 100000;
+    //vert1z[Z] /= 100000;
+    //vert2z[Z] /= 100000;
+    //pointz[Z] /= 100000;
 
     GzCoord v0, v1, v2;
     v0[X] = vert0z[X] - vert2z[X];
@@ -607,7 +607,7 @@ int findGeneralPlaneEq(const GzCoord* (&tri)[3], GzCoord &out, float &d) {
     return GZ_SUCCESS;
 }
 
-int sortVerticesCW(const GzCoord* (&tri)[3], GzCoord* (&normals)[3]) {
+int sortVerticesCW( const GzCoord* (&tri)[3], GzCoord* (&normals)[3]) {
 
     // determine top vert
     // sort by Y ascending
@@ -664,7 +664,7 @@ int sortVerticesCW(const GzCoord* (&tri)[3], GzCoord* (&normals)[3]) {
     return GZ_SUCCESS;
 }
 
-void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCoord *color) {
+void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCoord *color, const GzCoord &Ks, const GzCoord &Kd, const GzCoord &Ka) {
 
     GzCoord E;
     E[X] = pt[X] - render->camera.position[X];
@@ -727,17 +727,19 @@ void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCo
         sum_diff[Z] += render->lights[i].color[Z] * LdotN;
     }
 
-    sum_spec[X] *= render->Ks[X];
-    sum_spec[Y] *= render->Ks[Y];
-    sum_spec[Z] *= render->Ks[Z];
+    sum_spec[X] *= Ks[X];
+    sum_spec[Y] *= Ks[Y];
+    sum_spec[Z] *= Ks[Z];
 
-    sum_diff[X] *= render->Kd[X];
-    sum_diff[Y] *= render->Kd[Y];
[…]

### 2015-12-01 Code/rend.h

@@ -5,7 +5,7 @@
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
 #define	DEFAULT_IM_Z	(-200)//(-10.0)  /* world coords for image plane origin */
-#define	DEFAULT_IM_Y	(0.0)//(5.0)    /* default look-at point = 0,0,0 */
+#define	DEFAULT_IM_Y	(100.0)//(5.0)    /* default look-at point = 0,0,0 */
 #define	DEFAULT_IM_X	(0.0)//(-10.0)
 
 #define	DEFAULT_AMBIENT	{0.1, 0.1, 0.1}
@@ -101,4 +101,4 @@ int findGeneralPlaneEq(const GzCoord* (&tri)[3], GzCoord &out, float &d);
 
 void normalizeGzCoord(GzCoord &v);
 
-float* TracePath(GzRay ray, int depth, GzWorldSpaceTriangles *tris, GzRender *render,int i,int j); //Remove i and j
+void TracePath(GzRay ray, GzWorldSpaceTriangles *tris, int depth, GzRender* render,GzColor &retColor); 

### 2015-12-01 Code/rend.cpp
Fixed Phong Color.

@@ -1115,7 +1115,20 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 
 			//float br = 1.0f;
 			//for each triangle
-			/*for (int k = 0; k < tris->tris.size() ; k++)
+			float t_test = 0;
+			float t_testMin = 1000000000.0f;
+			GzCoord triA_min;
+			GzCoord triB_min;
+			GzCoord triC_min;
+			const GzCoord* tri_w_min[3] = {};
+			GzCoord* tri_n_min[3] = {};
+			bool intersected = false;
+			GzCoord InterKs;
+			GzCoord InterKd;
+			GzCoord InterKa;
+
+			//Find nearest object intersection
+			/*for (int k = 0; k < tris->tris.size(); k++)
 			{
 				GzCoord triA;
 				triA[0] = tris->tris[k]->vertices[0]->pos[0];
@@ -1157,54 +1170,54 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 				tri_n[1] = &triBNorm;
 				tri_n[2] = &triCNorm;
 
-				sortVerticesCW(tri_w, tri_n);
-
-				float t_test = 0;
+				//sortVerticesCW(tri_w, tri_n);
+				t_test = 0;
 				bool test = rayTriangleIntersect(ray.position, ray.direction, triA, triB, triC, t_test);
 
 				if (test)
 				{
+					//Check min
+					if (t_test < t_testMin)
+					{
+						t_testMin = t_test;
+						//Phong Shading
+						GzCoord intersectionPt;
+						intersectionPt[0] = ray.position[0] + (ray.direction[0] * t_test);
+						intersectionPt[1] = ray.position[1] + (ray.direction[1] * t_test);
+						intersectionPt[2] = ray.position[2] + (ray.direction[2] * t_test);
 
-					//Phong Shading
-					GzCoord intersectionPt;
-					intersectionPt[0] = ray.position[0] + (ray.direction[0] * t_test);
-					intersectionPt[1] = ray.position[1] + (ray.direction[1] * t_test);
[…]

### 2015-12-01 Code/rend.h

@@ -4,8 +4,8 @@
 
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
-#define	DEFAULT_IM_Z	(-200)//(-10.0)  /* world coords for image plane origin */
-#define	DEFAULT_IM_Y	(100.0)//(5.0)    /* default look-at point = 0,0,0 */
+#define	DEFAULT_IM_Z	(200)//(-10.0)  /* world coords for image plane origin */
+#define	DEFAULT_IM_Y	(0.0)//(5.0)    /* default look-at point = 0,0,0 */
 #define	DEFAULT_IM_X	(0.0)//(-10.0)
 
 #define	DEFAULT_AMBIENT	{0.1, 0.1, 0.1}

### 2015-12-01 Code/Application5.cpp
Added Left Wall. Reflection Ray does not work.

@@ -123,10 +123,14 @@ GzMatrix	rotateY =
 	status |= GzBeginRender(m_pRender);
 
 	/* Light */
-	GzLight	light1 = { {-0.7071, 0.7071, 0}, {0.5, 0.5, 0.9} };
-	GzLight	light2 = { {0, -0.7071, -0.7071}, {0.9, 0.2, 0.3} };
-	GzLight	light3 = { {0.7071, 0.0, -0.7071}, {0.2, 0.7, 0.3} };
-	GzLight	ambientlight = { {0, 0, 0}, {0.3, 0.3, 0.3} };
+	GzLight	light1 = { { -0.7071, 0.7071, 0 }, { 0.9, 0.9, 0.9 } };
+	GzLight	light2 = { { 0, -0.7071, -0.7071 }, { 0.9, 0.9, 0.9 } };
+	GzLight	light3 = { {0.7071, 0.0, -0.7071}, {0.9, 0.9, 0.9} };
+	GzLight	ambientlight = { {0, 0, 0}, {0.9, 0.9, 0.9} };
+	//GzLight	light1 = { {-0.7071, 0.7071, 0}, {0.5, 0.5, 0.9} };
+	//GzLight	light2 = { {0, -0.7071, -0.7071}, {0.9, 0.2, 0.3} };
+	//GzLight	light3 = { {0.7071, 0.0, -0.7071}, {0.2, 0.7, 0.3} };
+	//GzLight	ambientlight = { {0, 0, 0}, {0.3, 0.3, 0.3} };
 
 	/* Material property */
 	GzColor specularCoefficient = { 0.3, 0.3, 0.3 };
@@ -146,7 +150,9 @@ GzMatrix	rotateY =
         valueListLights[1] = (GzPointer)&light2;
         nameListLights[2] = GZ_DIRECTIONAL_LIGHT;
         valueListLights[2] = (GzPointer)&light3;
-        status |= GzPutAttribute(m_pRender, 3, nameListLights, valueListLights);
+        //status |= GzPutAttribute(m_pRender, 1, nameListLights, valueListLights);
+		status |= GzPutAttribute(m_pRender, 3, nameListLights, valueListLights);
+
 
         nameListLights[0] = GZ_AMBIENT_LIGHT;
         valueListLights[0] = (GzPointer)&ambientlight;
@@ -237,9 +243,8 @@ int Application5::Render()
     // set kd,ks,ka (copied from above for now)
     /* Material property */
[…]

### 2015-12-01 Code/rend.cpp

@@ -1060,17 +1060,15 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			genColor[2] = 0;
 
 
-			for (int a = 0; a < 1; a++)
+			for (int a = 0; a < 4; a++)
 			{
 				GzColor c_;
 
 				if (a <= 0) //a < 0 for jitter
 				{
 					TracePath(ray, tris, 0, render, c_);
-					//c_[0] *= 4095;
-					//c_[1] *= 4095;
-					//c_[2] *= 4095;
-					GzPutDisplay(render->display, i, j, ctoi(c_[0]), ctoi(c_[1]), ctoi(c_[2]), 1.0f, 1.0f);
+
+					//GzPutDisplay(render->display, i, j, ctoi(c_[0]), ctoi(c_[1]), ctoi(c_[2]), 1.0f, 1.0f);
 
 
 				}
@@ -1091,9 +1089,7 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 					ray.position[1] = ray.position[1] + y_jitter;
 
 					TracePath(ray, tris, 0, render, c_);
-					//c_[0] *= 4095;
-					//c_[1] *= 4095;
-					//c_[2] *= 4095;
+
 				}
 				genColor[0] = genColor[0] + c_[0];
 				genColor[1] = genColor[1] + c_[1];
@@ -1103,15 +1099,15 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			}
 
 			//1/samples
-			//genColor[0] = genColor[0] * (.25); 
-			//genColor[1] = genColor[1] * (.25);
-			//genColor[2] = genColor[2] * (.25);
+			genColor[0] = genColor[0] * (.25); 
+			genColor[1] = genColor[1] * (.25);
+			genColor[2] = genColor[2] * (.25);
 
 			//USe for one sample
 			//GzColor c_;
 			//TracePath(ray, tris, 0, render, c_);
 			
-			//GzPutDisplay(render->display, i, j, ctoi(c_[0]), ctoi(c_[1]), ctoi(c_[2]), 1.0f, 1.0f);
+			GzPutDisplay(render->display, i, j, ctoi(genColor[0]), ctoi(genColor[1]), ctoi(genColor[2]), 1.0f, 1.0f);
 
 			//float br = 1.0f;
 			//for each triangle
[…]

### 2015-12-01 Code/rend.h

@@ -4,9 +4,9 @@
 
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
-#define	DEFAULT_IM_Z	(200)//(-10.0)  /* world coords for image plane origin */
-#define	DEFAULT_IM_Y	(0.0)//(5.0)    /* default look-at point = 0,0,0 */
-#define	DEFAULT_IM_X	(0.0)//(-10.0)
+#define	DEFAULT_IM_Z	(-200)//(-10.0)  /* world coords for image plane origin */
+#define	DEFAULT_IM_Y	(100.0)//(5.0)    /* default look-at point = 0,0,0 */
+#define	DEFAULT_IM_X	(10.0)//(-10.0)
 
 #define	DEFAULT_AMBIENT	{0.1, 0.1, 0.1}
 #define	DEFAULT_DIFFUSE	{0.7, 0.6, 0.5}

### 2015-12-02 paper/paper.tex
clarifications & fixes to paper

@@
 Monte Carlo path tracing combines the techniques of reflected or refracted rays by recursively generating other rays, but also captures the lighting relations between diffuse surfaces.
-Using random numbers, it simulates the integration found the rendering equation by using the Russian Roulette technique for the generation of every new ray, i.e.
+Using random numbers, it approximates the integration found the rendering equation by using the Russian Roulette technique for the generation of every new ray, i.e.
 a random number between zero and one is generated and compared to its colour coefficients.
@@
   \item[2)] Find the point of intersection with the nearest surface
-  \item[3)] Randomly (Russian Roulette) decide whether the light is emitted or reflected (test against Kd or Ks)
-  \item[3A)] if emitted, return weight * Le
-  \item[3B)] if refleced, weight *= Ks, randomly scatter the ray according to its BRDF pdf, and go to step 2
+  \item[3)] Randomly (Russian Roulette) decide whether the light is emitted or reflected (test against this random number against the Kd of the surface):
+  \item[3A)] if emitted, return weight * shading from light source(s)
+  \item[3B)] if reflected, weight *= Ks, randomly scatter the ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
 \end{description}
@@
 In order for this algorithm to be implemented, other algorithms such as ray-triangle intersection and Phong shading (for emitted light) must also be implemented. 
-Additionally, a BRDF probability distribution function (pdf) must be created for each surface. A reasonable BRDF pdf would be a uniform distribution, wherein \emph{any} 
+Additionally, a BRDF pdf must be created for each surface. A reasonable BRDF pdf would be a uniform distribution, wherein \emph{any} 
 vector (x,y,z) which is in the hemisphere of the surface is randomly selected.

### 2015-12-02 Code/Application5.cpp
Need to fix reflected ray and Test on the teapot.

@@ -242,14 +242,20 @@ int Application5::Render()
 
     // set kd,ks,ka (copied from above for now)
     /* Material property */
-	GzColor specularCoefficient = { 0.3, 0.3, 0.3 }; //Using for color
+	GzColor specularCoefficient = { 0.3, 0.3, 0.3 }; 
     GzColor ambientCoefficient = { 0.8, 0.1, 0.1 };
-    GzColor diffuseCoefficient = { 0.7, 0.7, 0.7 }; //Using for emission
+    GzColor diffuseCoefficient = { 0.7, 0.7, 0.7 }; 
+	GzColor emissionCoefficient = { 0.7, 0.7, 0.7 };
+
+	int left_wall = 2;
+	int right_wall = 5;
+	int back_wall = 8;
+	int floor_wall = 11;
+	int ceil_wall = 14;
+	int light_plane = 17;
+	int front_wall = 20;
 
 
-	int left_wall = 1;
-	int right_wall = 3;
-	int back_wall = 2;
 
 	int counter = 0;
 	while( fscanf(infile, "%s", dummy) == 1) { 	/* read in tri word */
@@ -278,28 +284,120 @@ int Application5::Render()
 			ambientCoefficient[0] = 0;
 			ambientCoefficient[1] = 1;
 			ambientCoefficient[2] = 0;
+
+			diffuseCoefficient[0] = 0;
+			diffuseCoefficient[1] = 1;
+			diffuseCoefficient[2] = 0;
+
+			emissionCoefficient[0] = 0.0f;
+			emissionCoefficient[1] = 0.0f;
+			emissionCoefficient[2] = 0.0f;
+
 			
 		}
-		else
+		else if (counter <= right_wall)
 		{
-			ambientCoefficient[0] = 1;
+			ambientCoefficient[0] = 0;
 			ambientCoefficient[1] = 0;
-			ambientCoefficient[2] = 0;
-		}
-		/*else if (counter <= right_wall)
-		{
-			specularCoefficient[0] = 0;
-			specularCoefficient[1] = 0;
-			specularCoefficient[2] = 1;
+			ambientCoefficient[2] = 1;
+
+			emissionCoefficient[0] = 0.0f;
+			emissionCoefficient[1] = 0.0f;
+			emissionCoefficient[2] = 0.0f;
 
 			diffuseCoefficient[0] = 0;
 			diffuseCoefficient[1] = 0;
[…]

### 2015-12-02 Code/disp.cpp

@@ -54,7 +54,7 @@ int GzInitDisplay(GzDisplay	*display)
 /* HW1.5 set everything to some default values - start a new frame */
 	for (int i = 0; i < display->xres; i++) {
 		for (int j = 0; j < display->yres; j++) {
-			display->fbuf[i + display->xres*j] = GzPixel{ GREY, 0, GREY, 1, INT_MAX };
+			display->fbuf[i + display->xres*j] = GzPixel{ 0.0, 0.0, 0.0, 1, INT_MAX };
 		}
 	}
 	return GZ_SUCCESS;

### 2015-12-02 Code/rend.cpp

@@ -994,7 +994,7 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			ray.position[1] = render->camera.position[1];
 			ray.position[2] = render->camera.position[2];
 
-			float d = 50;//1 / (tan(degToRad(render->camera.FOV) / 2.0f));
+			float d = 20;//1 / (tan(degToRad(render->camera.FOV) / 2.0f));
 			float W = 2;//2 * d*(tan(degToRad(render->camera.FOV) / 2.0f)); //Horizontal Field of View
 			float H = 2;// 2 * d*(tan(degToRad(render->camera.FOV) / 2.0f)); //Assuming same field of view in vertical
 			
@@ -1049,27 +1049,21 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			ray.direction[2] = pixelPos[2];
 			normalizeGzCoord(ray.direction);
 
-			//Recursive call!
 
-			//One Sample
-
-			//Four Samples
 			GzColor genColor;
 			genColor[0] = 0;
 			genColor[1] = 0;
 			genColor[2] = 0;
 
-
-			for (int a = 0; a < 4; a++)
+			//Sampling
+			int numSamples = 200;
+			for (int a = 0; a < numSamples; a++)
 			{
 				GzColor c_;
 
-				if (a <= 0) //a < 0 for jitter
+				if (a <= 0)
 				{
 					TracePath(ray, tris, 0, render, c_);
-
-					//GzPutDisplay(render->display, i, j, ctoi(c_[0]), ctoi(c_[1]), ctoi(c_[2]), 1.0f, 1.0f);
-
 
 				}
 				else
@@ -1099,124 +1093,12 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			}
 
 			//1/samples
-			genColor[0] = genColor[0] * (.25); 
-			genColor[1] = genColor[1] * (.25);
-			genColor[2] = genColor[2] * (.25);
+			genColor[0] = genColor[0] * (1 / (float)numSamples);
+			genColor[1] = genColor[1] * (1 / (float)numSamples);
+			genColor[2] = genColor[2] * (1 / (float)numSamples);
 
-			//USe for one sample
-			//GzColor c_;
[…]

### 2015-12-02 Code/rend.h

@@ -4,9 +4,9 @@
 
 /* Camera defaults */
 #define	DEFAULT_FOV		35.0
-#define	DEFAULT_IM_Z	(-200)//(-10.0)  /* world coords for image plane origin */
-#define	DEFAULT_IM_Y	(100.0)//(5.0)    /* default look-at point = 0,0,0 */
-#define	DEFAULT_IM_X	(10.0)//(-10.0)
+#define	DEFAULT_IM_Z	(-100)//(-10.0)  /* world coords for image plane origin */
+#define	DEFAULT_IM_Y	(0.0)//(5.0)    /* default look-at point = 0,0,0 */
+#define	DEFAULT_IM_X	(0.0)//(-10.0)
 
 #define	DEFAULT_AMBIENT	{0.1, 0.1, 0.1}
 #define	DEFAULT_DIFFUSE	{0.7, 0.6, 0.5}
@@ -41,7 +41,7 @@ typedef struct {			/* define a renderer */
 typedef struct {
     GzCoord pos;
     GzCoord normal;
-    GzColor Ka, Kd, Ks;
+    GzColor Ka, Kd, Ks, Le;
 } GzVertex;
 
 typedef struct {

### 2015-12-02 paper/paper.tex

@@
 Rendering via the use of rays, such as in the Monte Carlo path tracing approach, stands in stark contrast to the rasterization of triangles, which is the faster and more common approach to rendering, and the one used throughout the assignments of the CSCI 580 course.
-However, with path tracing it is possible to achieve photorealism by simulating the travel of photons, rather than the approximations used in renderers that make use of rasterization.
-INSERT STUFF RELATING TO WHAT WE ACCOMPLISHED.
+However, with path tracing it is possible to achieve photorealism by simulating the travel of light, which bounces off surfaces, resulting in shadows, specular as well as diffuse reflections andrefractions rather than the surface shading approach used in renderers that make use of rasterization.
 
@@
 \begin{description}
-  \item[1)] Generate a ray from the camera (position x,y, horizontal and vertical fields of view u,v), passing through the image plane, with weight = 1
+  \item[1)] Generate a ray from the camera (position x,y, horizontal and vertical fields of view u,v), passing through the image plane, with $weight = 1$
   \item[2)] Find the point of intersection with the nearest surface
   \item[3)] Randomly (Russian Roulette) decide whether the light is emitted or reflected (test against this random number against the Kd of the surface):
-  \item[3A)] if emitted, return weight * shading from light source(s)
-  \item[3B)] if reflected, weight *= Ks, randomly scatter the ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
+  \item[3A)] if emitted, return $weight * shading$ from light source(s)
+  \item[3B)] if reflected, $weight \mathrel{*}= Ks$, randomly scatter the ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
 \end{description}
@@
 \section{Implementation}
-Explain architectural changes.
-Explain what parts of Monte Carlo we did.
+Firstly, we used the codebase of the previous assignments as a launching point for the project.
+Thus, the GUI and initialization, as well other elements such as the rendering parameters, the functions for displaying pixels to the screen and a number of vector and matrix functions were borrowed directly from previous assignments.
+However, rasterization-related functionality was thrown out and replaced with an entirely new way of rendering via rays.
+\par
+The graphics pipeline was designed as follows: initialization is carried out, wherein global properties such as camera position and lights are defined.
+Then, the meshes are loaded into memory as a data structure encompassing all triangles.
+Each vertex in the triangle has a position, a normal, as well as information on their specular, diffuse and ambient coefficients, which differs from the global Ks, Kd and Ka found in the assignment renderer.
+
+\par
+asda
 

### 2015-12-02 paper/paper.tex
works continues on paper...

@@
 
-\section{Implementation}
+\section{Application Structure}
 Firstly, we used the codebase of the previous assignments as a launching point for the project.
@@
 However, rasterization-related functionality was thrown out and replaced with an entirely new way of rendering via rays.
-\par
+
+\section{Graphics Pipeline}
 The graphics pipeline was designed as follows: initialization is carried out, wherein global properties such as camera position and lights are defined.
@@
 Each vertex in the triangle has a position, a normal, as well as information on their specular, diffuse and ambient coefficients, which differs from the global Ks, Kd and Ka found in the assignment renderer.
+Next, this set of all geometry is used by the path tracing technique to determine colour at each pixel, which is finally fed to the same display as the one used in previous asignments.
 
-\par
-asda
+\section{The Path Tracing Implementation}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{raycast_viewplane}
+\caption{Casting a ray through the image plane}
+\label{raycast_viewplane}
+
+\end{figure}
 
@@
 Pictures, maybe some figures for different numbers of rays.
+
+\section{Challenges \& Observations}
 

### 2015-12-02 paper/paper.tex
more of the paper

@@
 Then, the meshes are loaded into memory as a data structure encompassing all triangles.
-Each vertex in the triangle has a position, a normal, as well as information on their specular, diffuse and ambient coefficients, which differs from the global Ks, Kd and Ka found in the assignment renderer.
+Each vertex in the triangle has a position and a normal in world space coordinates, as well as information on their specular, diffuse and ambient coefficients, which differs from the global Ks, Kd and Ka found in the assignment renderer.
 Next, this set of all geometry is used by the path tracing technique to determine colour at each pixel, which is finally fed to the same display as the one used in previous asignments.
@@
 \section{The Path Tracing Implementation}
+Much like the Monte Carlo path tracing algorithm outlined previously, rays are cast from the camera out into the scene, through the view plane.
+The reason for why the process is carried out in the reverse order in comparison to the real world, in which light is emitted from a light source, bounces off surfaces and hits the film or photoreceptors (or the real-world ``image plane'') of the camera, is that in reality, the vast majority of light misses this ``image plane''.
+Therefore, it would be extremely inefficient to calculate the path of light beams (or photons) which would be unseen by the camera.
+\par
+If we know the vertical and horizontal fields of view of the camera, as well as its position (e, for eye) and coordinate system (u,v,n) as well as picking an arbitrary distance (d) for the view plane, we can determine the height and width of the image plane:
+
+\begin{center}
+$\tan(\theta_{vertical}/2) = \text{height} / 2d$
+\end{center}
+
+\begin{flushright}
+\par 
+(2)
+\end{flushright}
+
+\begin{center}
+$\tan(\theta_{horizontal}/2) = \text{width} / 2d$
+\end{center}
+
+\begin{flushright}
+\par 
+(3)
+\end{flushright}
+
+Then, we can find the center of the image plane, and its bottom-left corner:
+
+\begin{center}
+$c = e - n * d$
+\end{center}
+
+\begin{flushright}
+\par 
+(4)
+\end{flushright}
+
+\begin{center}
+$\text{bottom left} = c - u * \text{height} /2 - v * \text{width} /2$
+\end{center}
+
+\begin{flushright}
+\par 
+(4)
+\end{flushright}
 
@@
 \includegraphics[width=2.5in]{raycast_viewplane}
-\caption{Casting a ray through the image plane}
-\label{raycast_viewplane}
+\caption{Casting a ray through the image plane[5]}
+\label{raycast_imageplane}
 
@@
 
+\bibitem {Chapter}
+\url{http://www.cs.virginia.edu/~gfx/Courses/2010/IntroGraphics/Lectures/6-RayCasting.pdf} slide 58
+
 \end{thebibliography}

### 2015-12-03 Code/Application5.cpp
Added Time. Still Crashes after 100 rays??

@@ -12,6 +12,7 @@
 #include "Gz.h"
 #include "disp.h"
 #include "rend.h"
+#include "time.h"
 
 #ifdef _DEBUG
 #undef THIS_FILE
@@ -21,7 +22,6 @@ static char THIS_FILE[]=__FILE__;
 
 #define INFILE  "ppot.asc"
 #define OUTFILE "output.ppm"
-
 
 extern int tex_fun(float u, float v, GzColor color); /* image texture function */
 extern int ptex_fun(float u, float v, GzColor color); /* procedural texture function */
@@ -127,11 +127,6 @@ GzMatrix	rotateY =
 	GzLight	light2 = { { 0, -0.7071, -0.7071 }, { 0.9, 0.9, 0.9 } };
 	GzLight	light3 = { {0.7071, 0.0, -0.7071}, {0.9, 0.9, 0.9} };
 	GzLight	ambientlight = { {0, 0, 0}, {0.9, 0.9, 0.9} };
-	//GzLight	light1 = { {-0.7071, 0.7071, 0}, {0.5, 0.5, 0.9} };
-	//GzLight	light2 = { {0, -0.7071, -0.7071}, {0.9, 0.2, 0.3} };
-	//GzLight	light3 = { {0.7071, 0.0, -0.7071}, {0.2, 0.7, 0.3} };
-	//GzLight	ambientlight = { {0, 0, 0}, {0.3, 0.3, 0.3} };
-
 	/* Material property */
 	GzColor specularCoefficient = { 0.3, 0.3, 0.3 };
 	GzColor ambientCoefficient = { 0.1, 0.1, 0.1 };
@@ -209,6 +204,8 @@ int Application5::Render()
 	char		dummy[256]; 
 	int			status; 
 
+	time_t start, stop;
+	time(&start);
 
 	/* Initialize Display */
 	status |= GzInitDisplay(m_pDisplay); 
@@ -285,9 +282,9 @@ int Application5::Render()
 			ambientCoefficient[1] = 1;
 			ambientCoefficient[2] = 0;
 
-			diffuseCoefficient[0] = 0;
-			diffuseCoefficient[1] = 1;
-			diffuseCoefficient[2] = 0;
+			diffuseCoefficient[0] = .25;
+			diffuseCoefficient[1] = .75;
+			diffuseCoefficient[2] = .25;
 
 			emissionCoefficient[0] = 0.0f;
 			emissionCoefficient[1] = 0.0f;
@@ -301,55 +298,55 @@ int Application5::Render()
[…]

### 2015-12-03 Code/rend.cpp

@@ -10,7 +10,8 @@
 
 #define PI 3.141592653589793f
 #define MAXREFLECTANCE 10
-
+const int RAY_SAMPLES = 100;
+const int RAY_DEPTH = 6;
 // HW3 helpers
 float degToRad(float deg) {
     return deg * PI / 180.0f;
@@ -1056,19 +1057,19 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			genColor[2] = 0;
 
 			//Sampling
-			int numSamples = 200;
+			int numSamples = RAY_SAMPLES;
 			for (int a = 0; a < numSamples; a++)
 			{
 				GzColor c_;
 
-				if (a <= 0)
-				{
+				//if (a <= 0)
+				//{
 					TracePath(ray, tris, 0, render, c_);
 
-				}
-				else
-				{
-					float random = ((float)rand()) / (float)RAND_MAX;
+				//}
+				//else
+				//{
+					/*float random = ((float)rand()) / (float)RAND_MAX;
 					float diff = 1.0f - 0.0f;
 					float r = random * diff;
 					float ran_num = r;
@@ -1080,14 +1081,14 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 					float y_jitter = (ran_num*m_y_spacing) - (m_y_spacing / .5f);
 
 					ray.position[0] = ray.position[0] + x_jitter;
-					ray.position[1] = ray.position[1] + y_jitter;
+					ray.position[1] = ray.position[1] + y_jitter;*/
 
-					TracePath(ray, tris, 0, render, c_);
+				//	TracePath(ray, tris, 0, render, c_);
 
-				}
-				genColor[0] = genColor[0] + c_[0];
-				genColor[1] = genColor[1] + c_[1];
-				genColor[2] = genColor[2] + c_[2];
+	//			}
+				genColor[0] += c_[0];
+				genColor[1] += c_[1];
+				genColor[2] += c_[2];
 
 
 			}
@@ -1236,17 +1237,17 @@ void TracePath(GzRay ray, GzWorldSpaceTriangles *tris, int depth, GzRender *rend
 		
 
[…]

### 2015-12-03 paper/paper.tex
paper continues..

@@
 \par
-If we know the vertical and horizontal fields of view of the camera, as well as its position (e, for eye) and coordinate system (u,v,n) as well as picking an arbitrary distance (d) for the view plane, we can determine the height and width of the image plane:
+If we know the vertical and horizontal fields of view of the camera, as well as its position (e, for eye) and coordinate system (u,v,n) as well as picking an arbitrary distance (d) for the view plane, we can determine the height ($h$) and width ($w$) of the image plane:
 
 \begin{center}
-$\tan(\theta_{vertical}/2) = \text{height} / 2d$
+$\tan(\theta_{vertical}/2) = h / 2d$
 \end{center}
@@
 \begin{center}
-$\tan(\theta_{horizontal}/2) = \text{width} / 2d$
+$\tan(\theta_{horizontal}/2) =w / 2d$
 \end{center}
@@
 
-Then, we can find the center of the image plane, and its bottom-left corner:
+Then, we can find the center of the image plane, and its bottom-left corner ($BL$):
 
@@
 \begin{center}
-$\text{bottom left} = c - u * \text{height} /2 - v * \text{width} /2$
+$BL = c - u * h /2 - v *w/2$
 \end{center}
@@
 \par 
-(4)
+(5)
+\end{flushright}
+
+Finally, with knowledge of the image resolution ($w_{i}$ and $h_{i}$), we can find the location of the pixel that the ray will pass through:
+
+\begin{center}
+$\text{pixel}_{(i,j)}= BL + u * i * w / w_{i} + v * j * h/h_{i}$
+\end{center}
+
+\begin{flushright}
+\par 
+(6)
 \end{flushright}

### 2015-12-03 Code/Application5.cpp
Added Cube. Screw the Teapot.

@@ -76,31 +76,6 @@ int Application5::Initialize()
  
 	status |= GzNewRender(&m_pRender, m_pDisplay); 
 
-/* Translation matrix */
-GzMatrix	scale = 
-{ 
-	3.25,	0.0,	0.0,	0.0, 
-	0.0,	3.25,	0.0,	-3.25, 
-	0.0,	0.0,	3.25,	3.5, 
-	0.0,	0.0,	0.0,	1.0 
-}; 
- 
-GzMatrix	rotateX = 
-{ 
-	1.0,	0.0,	0.0,	0.0, 
-	0.0,	.7071,	.7071,	0.0, 
-	0.0,	-.7071,	.7071,	0.0, 
-	0.0,	0.0,	0.0,	1.0 
-}; 
- 
-GzMatrix	rotateY = 
-{ 
-	.866,	0.0,	-0.5,	0.0, 
-	0.0,	1.0,	0.0,	0.0, 
-	0.5,	0.0,	.866,	0.0, 
-	0.0,	0.0,	0.0,	1.0 
-}; 
-
 #if 0 	/* set up app-defined camera if desired, else use camera defaults */
     camera.position[X] = -3;
     camera.position[Y] = -25;
@@ -174,19 +149,9 @@ GzMatrix	rotateY =
         specpower = 32;
         valueListShader[4] = (GzPointer)&specpower;
 
-        // get rid of textures
-        //nameListShader[5]  = GZ_TEXTURE_MAP;
-#if 0   /* set up null texture function or valid pointer */
-        //valueListShader[5] = (GzPointer)0;
-#else
-        //valueListShader[5] = (GzPointer)(tex_fun);	/* or use ptex_fun */
-#endif
+
         status |= GzPutAttribute(m_pRender, 5, nameListShader, valueListShader);
 
-
-	//status |= GzPushMatrix(m_pRender, scale);  
-	//status |= GzPushMatrix(m_pRender, rotateY); 
-	//status |= GzPushMatrix(m_pRender, rotateX); 
 
 	if (status) exit(GZ_FAILURE); 
 
@@ -376,27 +341,24 @@ int Application5::Render()
 			diffuseCoefficient[1] = .75;
 			diffuseCoefficient[2] = .75;
 		}
-		else //TeaPot
+		else //Cube
 		{
-			ambientCoefficient[0] = .7;
-			ambientCoefficient[1] = .1;
-			ambientCoefficient[2] = .1;
+			ambientCoefficient[0] = .4;
[…]

### 2015-12-03 Code/rend.cpp

@@ -10,8 +10,8 @@
 
 #define PI 3.141592653589793f
 #define MAXREFLECTANCE 10
-const int RAY_SAMPLES = 100;
-const int RAY_DEPTH = 6;
+const int RAY_SAMPLES = 1;
+const int RAY_DEPTH = 1;
 // HW3 helpers
 float degToRad(float deg) {
     return deg * PI / 180.0f;
@@ -86,11 +86,6 @@ void barycentricCoords(const GzCoord &vert0, const GzCoord &vert1, const GzCoord
     memcpy(vert2z, vert2, sizeof(GzCoord));
     memcpy(pointz, point, sizeof(GzCoord));
 
-    //reduce z range to prevent overflow
-    //vert0z[Z] /= 100000;
-    //vert1z[Z] /= 100000;
-    //vert2z[Z] /= 100000;
-    //pointz[Z] /= 100000;
 
     GzCoord v0, v1, v2;
     v0[X] = vert0z[X] - vert2z[X];
@@ -1062,30 +1057,8 @@ int raycast_render(GzRender *render, GzWorldSpaceTriangles *tris)
 			{
 				GzColor c_;
 
-				//if (a <= 0)
-				//{
-					TracePath(ray, tris, 0, render, c_);
+				TracePath(ray, tris, 0, render, c_);
 
-				//}
-				//else
-				//{
-					/*float random = ((float)rand()) / (float)RAND_MAX;
-					float diff = 1.0f - 0.0f;
-					float r = random * diff;
-					float ran_num = r;
-					float x_jitter = (ran_num*m_x_spacing) - (m_x_spacing / .5f);
-
-					random = ((float)rand()) / (float)RAND_MAX;
-					r = random * diff;
-					ran_num = r;
-					float y_jitter = (ran_num*m_y_spacing) - (m_y_spacing / .5f);
-
-					ray.position[0] = ray.position[0] + x_jitter;
-					ray.position[1] = ray.position[1] + y_jitter;*/
-
-				//	TracePath(ray, tris, 0, render, c_);
-
-	//			}
 				genColor[0] += c_[0];
 				genColor[1] += c_[1];
 				genColor[2] += c_[2];
@@ -1217,16 +1190,14 @@ void TracePath(GzRay ray, GzWorldSpaceTriangles *tris, int depth, GzRender *rend
 	
[…]

### 2015-12-03 paper/paper.tex
more sections of paper

@@
 Monte Carlo path tracing combines the techniques of reflected or refracted rays by recursively generating other rays, but also captures the lighting relations between diffuse surfaces.
-Using random numbers, it approximates the integration found the rendering equation by using the Russian Roulette technique for the generation of every new ray, i.e.
-a random number between zero and one is generated and compared to its colour coefficients.
+We implemented it with ray reflections in a random uniform fashion, with a deterministic number of reflected rays or until the ray hits a light source.
 
 \section{The Monte Carlo Path Tracing Algorithm}
-The essentials of the Monte Carlo algorithm are surprisingly simple and can be summed as thus[4]:
+The core of the Monte Carlo algorithm is recursive and has relatively few steps. It can be summed as thus(modified version of an algorithm found in [4]):
 
@@
   \item[2)] Find the point of intersection with the nearest surface
-  \item[3)] Randomly (Russian Roulette) decide whether the light is emitted or reflected (test against this random number against the Kd of the surface):
-  \item[3A)] if emitted, return $weight * shading$ from light source(s)
-  \item[3B)] if reflected, $weight \mathrel{*}= Ks$, randomly scatter the ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
+  \item[3)] Determine whether the surface is a light source or not (as indicated by a special coefficients for light emittance)
+  \item[3A)] if it is a light source, return $weight * light$
+  \item[3B)] if it is not, $weight \mathrel{*}= (Ks+Kd+Ka)/3$ and randomly scatter a new ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
 \end{description}
 
-In order for this algorithm to be implemented, other algorithms such as ray-triangle intersection and Phong shading (for emitted light) must also be implemented. 
+In order for this algorithm to be implemented, other algorithms such as ray-triangle intersection must also be implemented. 
 Additionally, a BRDF pdf must be created for each surface. A reasonable BRDF pdf would be a uniform distribution, wherein \emph{any} 
@@
 Each vertex in the triangle has a position and a normal in world space coordinates, as well as information on their specular, diffuse and ambient coefficients, which differs from the global Ks, Kd and Ka found in the assignment renderer.
+Also associated with vertices is a lighting coefficient for surfaces which emit light.
 Next, this set of all geometry is used by the path tracing technique to determine colour at each pixel, which is finally fed to the same display as the one used in previous asignments.
@@
 
+Thus, with the position of the pixel we want to draw on the image plane, and the position of the camera, we can draw a ray starting at the camera, passing through that pixel, and into the environment.
+
 \begin{flushright}
@@
 
+\par
+Once the ray is defined, the closest triangle (in z) that intersects with the ray is found. This is done in a brute-force approach, iterating through all triangles.
+Next, as outlined in the Monte Carlo path tracing algorithm, we recursively reflect the ray up to a set maximum amount of rays (depth), unless the ray hits a light source, in which case the lighting is returned.
+If no light source is found with the maximum depth, the pixel is deemed to have received no light and is shaded as black.
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{shad2-indirectdiffusehemi.png}
+\caption{Random uniform creation of rays in the hemisphere of the collision[6]}
+\label{hemisphere rays}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{recursive_rays}
+\caption{Recursively casting rays between surfaces from the camera to a light source[5]}
+\label{recursive_rays}
+
+\end{figure}
+
+\par
+For each pixel for which the above process is carried out, it is repeated a certain number of times (or samples), so as to obtain an average which converges to the expected value of lighting for that pixel.
+
 \section{Results}
@@
 
+\section{Further Work}
+Because of the limited scope of the project, many features that could be implemented using Monte Carlo path tracing are missing.
+For example, there is currently no difference between a highly specular and highly diffuse surface because of the uniform random distribution of all reflected rays.
+Moreover, transparent surfaces and refraction are not simulated again due to the nature of the ray reflectance in the application.
+
 %===================================================
 
-\begin{thebibliography}{4}
+\begin{thebibliography}{5}
 
@@
 
-\bibitem {Chapter}
-\url{http://cs.brown.edu/courses/cs224/papers/mc_pathtracing.pdf} page 9
+\bibitem {Brown}
+\url{http://cs.brown.edu/courses/cs224/papers/mc_pathtracing.pdf}
 
-\bibitem {Chapter}
-\url{http://www.cs.virginia.edu/~gfx/Courses/2010/IntroGraphics/Lectures/6-RayCasting.pdf} slide 58
+\bibitem {Virginia}
+\url{http://www.cs.virginia.edu/~gfx/Courses/2010/IntroGraphics/Lectures/6-RayCasting.pdf}
+
+\bibitem {Scratchapixel}
+\url{http://www.scratchapixel.com/lessons/3d-basic-rendering/global-illumination-path-tracing/global-illumination-path-tracing-practical-implementation}
 

### 2015-12-03 Code/Application5.cpp
Free Bary! Added Images Folder.

@@ -354,9 +354,6 @@ int Application5::Render()
 			diffuseCoefficient[0] = .4;
 			diffuseCoefficient[1] = .3;
 			diffuseCoefficient[2] = .3;
-
-			
-
 		}
 
         GzTriangle* tri = new GzTriangle();

### 2015-12-03 Code/rend.cpp

@@ -10,8 +10,8 @@
 
 #define PI 3.141592653589793f
 #define MAXREFLECTANCE 10
-const int RAY_SAMPLES = 1;
-const int RAY_DEPTH = 1;
+const int RAY_SAMPLES = 40;
+const int RAY_DEPTH = 2;
 // HW3 helpers
 float degToRad(float deg) {
     return deg * PI / 180.0f;
@@ -1221,7 +1221,7 @@ void TracePath(GzRay ray, GzWorldSpaceTriangles *tris, int depth, GzRender *rend
 		pt_normal[Z] = (*bary)[X] * (tri_n_min[X])[Z] + (*bary)[Y] * (tri_n_min[Y])[Z] + (*bary)[Z] * (tri_n_min[Z])[Z];
 		//normalize in case of loss of precision
 		normalizeGzCoord(pt_normal);
-
+		free(bary);
 		GzCoord* pt_color = (GzCoord*)malloc(sizeof(GzCoord));
 		computeColorAtPt(render, point, pt_normal, pt_color, InterKs, InterKd, InterKa);
 

### 2015-12-03 Code/rend.cpp
Some optimizations. No Pointers.

@@ -10,8 +10,8 @@
 
 #define PI 3.141592653589793f
 #define MAXREFLECTANCE 10
-const int RAY_SAMPLES = 40;
-const int RAY_DEPTH = 2;
+const int RAY_SAMPLES = 10;
+const int RAY_DEPTH = 1;
 // HW3 helpers
 float degToRad(float deg) {
     return deg * PI / 180.0f;
@@ -660,7 +660,7 @@ int sortVerticesCW( const GzCoord* (&tri)[3], GzCoord* (&normals)[3]) {
     return GZ_SUCCESS;
 }
 
-void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCoord *color, const GzCoord &Ks, const GzCoord &Kd, const GzCoord &Ka) {
+void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCoord &color, const GzCoord &Ks, const GzCoord &Kd, const GzCoord &Ka) {
 
     GzCoord E;
     E[X] = pt[X] - render->camera.position[X];
@@ -731,9 +731,9 @@ void computeColorAtPt(GzRender *render, const GzCoord &pt, GzCoord &normal, GzCo
     sum_diff[Y] *= Kd[Y];
     sum_diff[Z] *= Kd[Z];
 
-    (*color)[X] = Ka[X] * render->ambientlight.color[X] + sum_diff[X] + sum_spec[X];
-    (*color)[Y] = Ka[Y] * render->ambientlight.color[Y] + sum_diff[Y] + sum_spec[Y];
-    (*color)[Z] = Ka[Z] * render->ambientlight.color[Z] + sum_diff[Z] + sum_spec[Z];
+    (color)[X] = Ka[X] * render->ambientlight.color[X] + sum_diff[X] + sum_spec[X];
+    (color)[Y] = Ka[Y] * render->ambientlight.color[Y] + sum_diff[Y] + sum_spec[Y];
+    (color)[Z] = Ka[Z] * render->ambientlight.color[Z] + sum_diff[Z] + sum_spec[Z];
 
 
 }
@@ -1095,6 +1095,7 @@ void TracePath(GzRay ray, GzWorldSpaceTriangles *tris, int depth, GzRender *rend
 	GzCoord InterKd;
 	GzCoord InterKa;
 	GzCoord InterLe;
+	int intersectIndex = 0;
 
[…]

### 2015-12-03 paper/paper.tex
paper commit...

@@
 \section{Results}
-Pictures, maybe some figures for different numbers of rays.
+
+To test our path tracing application, we placed a grey box floating in an environment that is similar to a typical Cornell Box, but with blue and green walls rather than red and green ones.
+When running the application with different values of $s$ (ray samples from each pixel) and $d$ (recursion depth), we found that the time of execution was linearly related to the amount of samples, which is unsurprising since $s$ controls the number of iterations of a single loop.
+When increasing $d$ however, the application appears to converge towards a given time.
+This can be explained by the geometry of the scene having an average number of reflections before hitting a light source.
+Therefore, beyond this value for d, the application does not take significantly more time to run.
+
+\par
+The image becomes sharp at around $s=10$ for any value of $d$, and artifacts persist beyond $s=40, d=4$.
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{1s_1d_26s}
+\caption{1 set of rays, recursion depth 1, time: 26 seconds}
+\label{1s_1d_26s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{1s_4d_34s}
+\caption{1 set of rays, recursion depth 4, time: 34 seconds}
+\label{1s_4d_34s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{1s_16d_36s}
+\caption{1 set of rays, recursion depth 16, time: 36 seconds}
+\label{1s_16d_36s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{10s_1d_5m42s}
+\caption{10 sets of rays, recursion depth 1, time: 5m42s}
+\label{10s_1d_5m42s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{20s_1d_11m26s}
+\caption{20 sets of rays, recursion depth 1, time: 11m26s}
+\label{20s_1d_11m26s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{40s_4d_23m25s}
+\caption{40 sets of rays, recursion depth 4, time: 23m25s}
+\label{40s_4d_23m25s}
+
+\end{figure}
 
@@
 \bibitem {Brown}
+Slides from Brown University.
 \url{http://cs.brown.edu/courses/cs224/papers/mc_pathtracing.pdf}
@@
 \bibitem {Virginia}
+Slides from the University of Virginia.
 \url{http://www.cs.virginia.edu/~gfx/Courses/2010/IntroGraphics/Lectures/6-RayCasting.pdf}
@@
 \bibitem {Scratchapixel}
+Global Illumination and Path Tracing.
 \url{http://www.scratchapixel.com/lessons/3d-basic-rendering/global-illumination-path-tracing/global-illumination-path-tracing-practical-implementation}

### 2015-12-03 paper/paper.tex
more work on paper

@@
 \section{Abstract}
-This will depend on our results.\\
-I predict this paper to be around 4 pages since it's in a rather condensed format.
+A Monte Carlo path tracing renderer was created using the assignment application as a base.
+Realistic images featuring shadows and diffuse interreflections become crisp beyond 20 sets of generated rays, with a recursion depth greater than or equal to 4.
 
@@
 \par
-The image becomes sharp at around $s=10$ for any value of $d$, and artifacts persist beyond $s=40, d=4$.
+The image becomes sharp at around $s=10$ for any value of $d$, and artifacts persist up to $s=40, d=4$ and $s=100,d=1$.
 
@@
 
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{100s_1d_40m21s}
+\caption{100 sets of rays, recursion depth 1, time: 40m21s}
+\label{100s_1d_40m21s}
+
+\end{figure}
+
 \section{Challenges \& Observations}
+There were many obstacles on the way to the completion of the project.
+Firstly, the codebase of the assignments is not particularly suited for ray casting: rather, it is suited for processing the geometry one triangle at a time since it rasterized them.
+Thus, a new architectured had to be written to replace the old one, which was more of a time sink than first anticipated.
+\par
+Second, building the scene was a long and tedious process, as the geometry had to be written in the ASCII format as read by the assignment renderer.
+Exacerbating this process is that path tracing is an expensive algorithm: to test the placement of geometry, the addition of features in the application or to debug it, one must wait between half a minute up to an hour depending on the quality of the generated image.
+\par
+Additionally, until a late phase of the project, we opted to continue to use the teapot that was featured in all previous assignments.
+However, since the teapot contains a great number of triangles, it was  expensive to calculate which triangle intersecting with the ray was the closest, since it required iterating through all the triangles.
+Since this operation was conducted per ray cast, the run-time cost of rendering the teapot was very high.
 
@@
 Moreover, transparent surfaces and refraction are not simulated again due to the nature of the ray reflectance in the application.
+
+\par
+Another opportunity for extension of the application is in its performance.
+There are two main avanues for this: improvement of runtime complexity and parallelization.
+In terms of runtime complexity, for determining the closest triangle intersected, one could make use of binary space partition trees rather than a brute force approach.
+Also, bounding volumes could be used to test for intersection with objects with complex geometries, such as the teapot, which would be encapsulated in a much more simple box shape for collision testing.
+
+\par
+As for parallelization, the ray casting approach lends itself well to being parallelized since each ray is reflected irrespective of rays casting through the same pixels (the samples), as well as irrespectively of the rays being cast through other pixels.
+Therefore, each individual ray could be calculated simultaneously.
+Incidentally, this process lends itself well to GPUs, which specialize in computing large sets of similar data.
 

### 2015-12-03 paper/paper.tex
paper finished (hopefully)

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[12pt,journal]{IEEEtran}
 
@@
   \item[3A)] if it is a light source, return $weight * light$
-  \item[3B)] if it is not, $weight \mathrel{*}= (Ks+Kd+Ka)/3$ and randomly scatter a new ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
+  \item[3B)] if not, $weight \mathrel{*}= \frac{Ks+Kd+Ka}{3}$ and randomly scatter a new ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
 \end{description}
+
+\newpage
 
@@
 
-Thus, with the position of the pixel we want to draw on the image plane, and the position of the camera, we can draw a ray starting at the camera, passing through that pixel, and into the environment.
-
 \begin{flushright}
@@
 \end{flushright}
+
+Thus, with the position of the pixel we want to draw on the image plane, and the position of the camera, we can draw a ray starting at the camera, passing through that pixel, and into the environment.
 
@@
 Therefore, beyond this value for d, the application does not take significantly more time to run.
+Additionally, an increase in $d$ also increases the brightness of the image, since a ray has a higher chance of hitting a light source if it is allowed a greater number of ``bounces'' off surfaces.
 
@@
 \centering
-\includegraphics[width=2.5in]{20s_1d_11m26s}
+\includegraphics[width=2.6in]{20s_1d_11m26s}
 \caption{20 sets of rays, recursion depth 1, time: 11m26s}
@@
 \label{100s_1d_40m21s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.6in]{100s_16d_42m10s}
+\caption{100 sets of rays, recursion depth 1, time: 42m10s}
+\label{1100s_16d_42m10s}
 
@@
 
+\par
+Also, as noted in the result sections, where an increase in $d$ increases brightness, the reflection quantity could be controlled by a random value to keep the shading consistent.
+This approach is called the Russian Roulette approach, wherein each successive reflected ray diminishes the chance of the next collision producing a reflected ray.
+This is done by comparing a random value between 0 and 1 to the current weight of the ray, which diminishes with each successive reflection.
+
+\par
+Furthermore, for images where $s \ge 10$, fully dark pixels are displayed in a noticeable, somewhat circular pattern.
+This could be the result of the uniform random reflectance along the semi-circular hemisphere, although there is currently not a definitive answer to why these artifacts appear.
+
 %===================================================

### 2015-12-03 paper/paper.tex
typo fix

@@
 Rendering via the use of rays, such as in the Monte Carlo path tracing approach, stands in stark contrast to the rasterization of triangles, which is the faster and more common approach to rendering, and the one used throughout the assignments of the CSCI 580 course.
-However, with path tracing it is possible to achieve photorealism by simulating the travel of light, which bounces off surfaces, resulting in shadows, specular as well as diffuse reflections andrefractions rather than the surface shading approach used in renderers that make use of rasterization.
+However, with path tracing it is possible to achieve photorealism by simulating the travel of light, which bounces off surfaces, resulting in shadows, specular as well as diffuse reflections and refractions rather than the surface shading approach used in renderers that make use of rasterization.
 

### 2015-12-03 paper/paper.tex
reorder paragraph + beginning of slides
finish powerpoint and talk about multithread in paper

@@
   \item[3A)] if it is a light source, return $weight * light$
-  \item[3B)] if not, $weight \mathrel{*}= \frac{Ks+Kd+Ka}{3}$ and randomly scatter a new ray according to a probability density function (pdf) approximating its bidirectional refletance distribution function (brdf), and go to step 2
+  \item[3B)] if not, $weight \mathrel{*}= \frac{Ks+Kd+Ka}{3}$ and randomly scatter a new ray according to a probability density function (pdf) approximating its bidirectional reflectance  distribution function (brdf), and go to step 2
 \end{description}
@@
 Also associated with vertices is a lighting coefficient for surfaces which emit light.
-Next, this set of all geometry is used by the path tracing technique to determine colour at each pixel, which is finally fed to the same display as the one used in previous asignments.
+Next, this set of all geometry is used by the path tracing technique to determine colour at each pixel, which is finally fed to the same display as the one used in previous assignments.
 
@@
 
+\par
+As for parallelization, the ray casting approach lends itself well to being parallelized since each ray is reflected irrespective of rays casting through the same pixels (the samples), as well as irrespectively of the rays being cast through other pixels.
+Therefore, multithreaded was implemented, where individual ray could be calculated simultaneously
+
 \section{Results}
@@
 \par
-The image becomes sharp at around $s=10$ for any value of $d$, and artifacts persist up to $s=40, d=4$ and $s=100,d=1$.
+Multithreading was mostly not used in the generation of the images, so that times could better be compared, although for high values of $s$ images were generated with multithreading.
+Since the application uses 25 threads, we see a decrease in speed by 25 times!
+
+\par
+The image becomes sharp at around $s=10$ for any value of $d$, but artifacts persist all the way up to $s=200$, $d=16$.
 
@@
 \centering
-\includegraphics[width=2.5in]{10s_1d_5m42s}
+\includegraphics[width=2.6in]{10s_1d_5m42s}
 \caption{10 sets of rays, recursion depth 1, time: 5m42s}
@@
 \centering
-\includegraphics[width=2.6in]{20s_1d_11m26s}
+\includegraphics[width=2.5in]{20s_1d_11m26s}
 \caption{20 sets of rays, recursion depth 1, time: 11m26s}
@@
 \centering
-\includegraphics[width=2.5in]{100s_1d_40m21s}
-\caption{100 sets of rays, recursion depth 1, time: 40m21s}
+\includegraphics[width=2.6in]{100s_1d_40m21s}
+\caption{100 sets of rays, recursion depth 1, time: 40m21s. Time with multithreading: 2min52s}
 \label{100s_1d_40m21s}
@@
 \centering
-\includegraphics[width=2.6in]{100s_16d_42m10s}
-\caption{100 sets of rays, recursion depth 1, time: 42m10s}
-\label{1100s_16d_42m10s}
+\includegraphics[width=2.5in]{100s_16d_42m10s}
+\caption{100 sets of rays, recursion depth 16, time: 42m10s. Time with multithreading: 2min59s}
+\label{100s_16d_42m10s}
+
+\end{figure}
+
+\begin{figure}[!t]
+
+\centering
+\includegraphics[width=2.5in]{200s_16d_5min53s}
+\caption{200 sets of rays, recursion depth 16, time with multithreading: 5min53s}
+\label{200s_16d_5min53s}
 
@@
 Firstly, the codebase of the assignments is not particularly suited for ray casting: rather, it is suited for processing the geometry one triangle at a time since it rasterized them.
-Thus, a new architectured had to be written to replace the old one, which was more of a time sink than first anticipated.
+Thus, a new architecture had to be written to replace the old one, which was more of a time sink than first anticipated.
 \par
@@
 Since this operation was conducted per ray cast, the run-time cost of rendering the teapot was very high.
+\par
+Furthermore, for images where $s \ge 10$, fully dark pixels are displayed in a noticeable, somewhat circular pattern.
+This could be the result of the uniform random reflectance along the semi-circular hemisphere, although there is currently not a definitive answer to why these artifacts appear.
 
@@
 Another opportunity for extension of the application is in its performance.
-There are two main avanues for this: improvement of runtime complexity and parallelization.
 In terms of runtime complexity, for determining the closest triangle intersected, one could make use of binary space partition trees rather than a brute force approach.
-Also, bounding volumes could be used to test for intersection with objects with complex geometries, such as the teapot, which would be encapsulated in a much more simple box shape for collision testing.
+Also, bounding volumes could be used to test for intersection with objects with complex geometries, such as the teapot, which would be encapsulated in a much simpler box shape for collision testing.
 
 \par
-As for parallelization, the ray casting approach lends itself well to being parallelized since each ray is reflected irrespective of rays casting through the same pixels (the samples), as well as irrespectively of the rays being cast through other pixels.
-Therefore, each individual ray could be calculated simultaneously.
-Incidentally, this process lends itself well to GPUs, which specialize in computing large sets of similar data.
+Although the application does use multithreading, the algorithm could also lend itself well to a GPU, which specializes in computing large sets of similar data.
 
@@
 This is done by comparing a random value between 0 and 1 to the current weight of the ray, which diminishes with each successive reflection.
-
-\par
-Furthermore, for images where $s \ge 10$, fully dark pixels are displayed in a noticeable, somewhat circular pattern.
-This could be the result of the uniform random reflectance along the semi-circular hemisphere, although there is currently not a definitive answer to why these artifacts appear.
 
